*Prochlorococcus marinus* responses to light and oxygen

### 2024-06-10 Code/Manuscript_O2.Rmd
added DOI
minor edit

Guided in part by the evidence of ocean distributions of proteins from
*P. marinus* we set up a matrix of photoperiods, PAR, and [O~2~] to approximate current, and potential future, latitudinal, depth and seasonal niches for *P. marinus* strains. As
mentioned, growth rate trials under red light, although not representative of *P. marinus* niches, are mechanistically informative [@murphyPhotoinactivationPhotosystemII2017] regarding photoinactivation of PSII. We implemented measures to minimize shock to cultures from exposure to experimental growth conditions by inoculating them the day before the experiment began, employing a sinusoidal photoperiod with a gradual increase in PAR exposure, and extracting growth rates from logistic curves fit over approximately 5 generations of growth, to accommodate multiple generations to acclimate to the imposed growth conditions [@mooreCulturingMarineCyanobacterium2007]. Although the current subtropical distribution of *P. marinus* spans a narrow range of photoperiods at the surface, light attenuation with depth shortens effective photoperiods as well as lowering peak PAR. Potential poleward
latitudinal range expansions, in combination with attenuation of light with depth, mean *P. marinus* clades may potentially encounter a wider range of photoperiods. Our growth rate determinations generally agree with those from Moore *et al*. [@mooreComparativePhysiologySynechococcus1995], for white LED and 250 µM O~2~, but our study is, to our knowledge, the first to analyze the interactive growth rate responses of *P. marinus* strains
to varying [O~2~], photoperiods and spectral wavebands.

@@

All data and metadata obtained from the Multicultivator were saved as comma separated
values files and are available on the Borealis data repository (DOI pending).

>>>>

Guided in part by the evidence of ocean distributions of proteins from
*P. marinus*, we set up a matrix of photoperiods, PAR, and [O~2~] to approximate current, and potential future, latitudinal, depth and seasonal niches for *P. marinus* strains. As
mentioned, growth rate trials under red light, although not representative of *P. marinus* niches, are mechanistically informative [@murphyPhotoinactivationPhotosystemII2017] regarding photoinactivation of PSII. We implemented measures to minimize shock to cultures from exposure to experimental growth conditions by inoculating them the day before the experiment began, employing a sinusoidal photoperiod with a gradual increase in PAR exposure, and extracting growth rates from logistic curves fit over approximately 5 generations of growth, to accommodate multiple generations to acclimate to the imposed growth conditions [@mooreCulturingMarineCyanobacterium2007]. Although the current subtropical distribution of *P. marinus* spans a narrow range of photoperiods at the surface, light attenuation with depth shortens effective photoperiods as well as lowering peak PAR. Potential poleward
latitudinal range expansions, in combination with attenuation of light with depth, mean *P. marinus* clades may potentially encounter a wider range of photoperiods. Our growth rate determinations generally agree with those from Moore *et al*. [@mooreComparativePhysiologySynechococcus1995], for white LED and 250 µM O~2~, but our study is, to our knowledge, the first to analyze the interactive growth rate responses of *P. marinus* strains
to varying [O~2~], photoperiods and spectral wavebands.

@@

All data are available on the figshare data repository (10.6084/m9.figshare.25959061).
